Supplement 6

### 2023-02-02 Code/Supplement 1B_figures.Rmd
all code and figures

---
title: "Supplement 1A"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

@@

library(scoringutils)
library(viridis)

library(ggeffects)
library(emmeans)
library(splines)

@@

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage
### Estimate Relative WIS 
pairWIS_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    distinct()
}
pairWIS_co_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model", "pop_size_quant2"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, pop_size_quant2, scaled_rel_skill) %>%
    distinct()
}

@@

## Function for setting up legends in empty facet spaces
library(gtable)
library(cowplot)

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}

@@

```{r Fig S1A, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=11}
start_date <-as.Date('2020-07-28')
stop_date <-as.Date('2021-12-21')

@@

## Must have forecasts for at least 50 locations for states/territories/district and national
include_locs <-function(x, df){

  df <-df %>%
    mutate(location=as.factor(location)) %>%
    group_by(model, forecast_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    filter(locs>=50)

@@

  df <-df %>%
    group_by(pop_size_quant2) %>%
    mutate(locs_total=length(unique(location))) %>% 
    ungroup() %>%
    group_by(model, pop_size_quant2, sub_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    mutate(percent_co=locs/locs_total) %>%
    filter(percent_co>=0.75)
}

@@

load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/updated_cases_nat_state_forecasts_2022-04-04.Rdata")

### National + state submissions
nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  filter(type=="point")

### National + state submissions that met inclusion criteria
teams_nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  include_quant(df=.) %>% #all quantiles present
  include_wks(df=.) %>% #all targets included 
  include_locs(df=.) %>% #at least 50 locations
  include_subs(df=.) %>% #submitted at least 50% of the time
  dplyr::select(model) %>%
  mutate(model_rename=paste0("*", model)) %>%
  distinct()

### Submission plots

nat_st %>%
  dplyr::select(-target, -target_end_date, -value) %>%
  distinct() %>%
  group_by(sub_date, model) %>%
  summarize(locs=n()) %>%
  ungroup() %>%
  mutate(log_loc=log(locs)) %>%
  ungroup() %>%
  mutate(total_sub=length(unique(sub_date))) %>%
  group_by(model) %>%
  mutate(wks_sub=length(sub_date),
         per_sub=wks_sub/total_sub) %>%
  ungroup() %>%
  left_join(., teams_nat_st, by="model") %>%
  mutate(model_rename=ifelse(is.na(model_rename), model, model_rename),
         model_rename=fct_reorder(model_rename, wks_sub)) %>%
  ggplot(., aes(sub_date, model_rename)) +
  geom_tile(aes(fill=log_loc)) +
  scale_fill_gradient(low = "#c7e9c0", high = "#41ab5d") + 
  geom_text(aes(label=locs), size=1.50) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B") +
  theme_bw() +
  labs(title = "A. Primary Model* Submissions for National/State/Territory-Level Case Forecasts Over Time",
       subtitle= "The number of locations forecasted per submission is labled",
       caption = "*Model met inclusion criteria") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        legend.position = "none")

@@

>>>>

---
title: "Supplement 1.B"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

@@

library(scoringutils)

@@

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage

@@

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=11}
start_date <-as.Date('2020-07-28')
stop_date <-as.Date('2021-12-21')

@@

@@

  df <-df %>%
    group_by(pop_size_quant2) %>%
    mutate(locs_total=length(unique(location))) %>% 
    ungroup() %>%
    group_by(model, pop_size_quant2, sub_date) %>%
    mutate(locs=length(unique(location))) %>% 
    ungroup() %>%
    mutate(percent_co=locs/locs_total) %>%
    filter(percent_co>=0.75)
}

@@

@@

```{r Fig S1B, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=11}

####
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/cases_county_forecasts_2022-04-04.Rdata")
size <-read.csv("C:/Users/oko8/Downloads/SVI2018_US_COUNTY.csv") %>%
  dplyr::select(FIPS, E_TOTPOP) %>%
  rename(location=FIPS) %>%
  mutate(location=as.character(location),
         location=ifelse(nchar(location)==4, paste0("0", location), location),
         pop_size_quant2=cut(E_TOTPOP, breaks = quantile(E_TOTPOP, probs = seq(0, 1, 0.20)),
                             include.lowest = TRUE, labels = 1:5)
  ) %>% 
  dplyr::select(-E_TOTPOP) 
all_dat <-all_dat %>% left_join(., size, by="location") 
rm(size)
counts <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  filter(type=="point")

# Inclusion criteria for counties based on pop size
all_dat <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date <= stop_date) 
all_dat <-all_dat %>% include_quant(df=.) 
all_dat <-all_dat %>% include_wks(df=.) 
all_dat <-all_dat %>% include_subs(df=.) 
all_dat <-all_dat %>% include_locs_co(df=.)

all_dat <-all_dat %>%
  dplyr::select(-quants, -no_horizons, -locs, -locs_total, -percent_co, -total, -subs, -percent_sub) %>%
  dplyr::select(model) %>%
  mutate(model_rename=paste0("*", model)) %>%
  distinct()

counts %>%
  dplyr::select(-target, -target_end_date, -value) %>%
  distinct() %>%
  group_by(sub_date, model, pop_size_quant2) %>%
  summarize(locs=n()) %>%
  rename_counts() %>%
  ungroup() %>%
  mutate(log_loc=log(locs)) %>%
  mutate(total_sub=length(unique(sub_date))) %>%
  ungroup() %>%
  group_by(model, pop_size_quant2) %>%
  mutate(wks_sub=length(sub_date),
         per_sub=wks_sub/total_sub) %>%
  ungroup() %>%
  left_join(., all_dat, by="model") %>%
  mutate(model_rename=ifelse(is.na(model_rename), model, model_rename),
         model_rename=fct_reorder(model_rename, wks_sub)) %>%
  ggplot(., aes(sub_date, model_rename)) +
  geom_tile(aes(fill=per_sub)) +
  scale_fill_gradient(low = "#c7e9c0", high = "#4a1486",
                      name="Percent of \nCounties \nForecasted") + 
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B") +
  facet_wrap(~ pop_size_quant2) +
  theme_bw() +
  labs(title = "B. Primary Model* Submissions for County-Level Case Forecasts Over Time",
       caption = "*Model met inclusion criteria") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        strip.background = strip.background_standard, strip.text = strip.text_standard,
        legend.position = "bottom")

### 2023-02-02 Code/Supplement 6_figures.Rmd
all code and figures

---
title: "Supplement 1A"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

@@

@@

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage
### Estimate Relative WIS 
pairWIS_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    distinct()
}
pairWIS_co_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model", "pop_size_quant2"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, pop_size_quant2, scaled_rel_skill) %>%
    distinct()
}

@@

## Change labels for counties
 rename_counts <-function(df) {   
   df <-df %>%
     mutate(pop_size_quant2=ifelse(pop_size_quant2==1, "County Forecasts: Size Quantile 1",
                                 ifelse(pop_size_quant2==2, "County Forecasts: Size Quantile 2",
                                        ifelse(pop_size_quant2==3, "County Forecasts: Size Quantile 3", 
                                               ifelse(pop_size_quant2==4, "County Forecasts: Size Quantile 4", "County forecasts: Size Quantile 5")))))
}

@@

### Supplement 1

```{r Fig S1A, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=11}
start_date <-as.Date('2020-07-28')
stop_date <-as.Date('2021-12-21')

sec_models <-c("CU-nochange",  "CU-scenario_high", "CU-scenario_low", "CU-scenario_mid", "COVIDhub_CDC-ensemble") ## Non-primary models should be exlcuded

## Must have 4 weeks of forecasts for all locations
include_wks <-function(x, df){

  df <-df %>%
    group_by(model, sub_date, location) %>%
    mutate(no_horizons=length(unique(target))) %>% 
    ungroup() %>%
    filter(no_horizons>=4) #There should be at least 4 horizons
}

## Must have quantities for all forecasts
include_quant <-function(x, df){

  df <-df %>%
    group_by(model, location, forecast_date, target) %>%
    mutate(quants=length(quantile[!is.na(quantile)])) %>%
    ungroup() %>%
    filter(quants==7) #There should be 7 quantiles

}

## Must have forecasts for at least 50 locations for states/territories/district and national
include_locs <-function(x, df){

  df <-df %>%
    mutate(location=as.factor(location)) %>%
    group_by(model, forecast_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    filter(locs>=50)

}

## Must have 75% of all counties per quantile per submission date forecasted
include_locs_co <-function(x, df){

  df <-df %>%
    group_by(pop_size_quant2) %>%
    mutate(locs_total=length(unique(location))) %>% 
    ungroup() %>%
    group_by(model, pop_size_quant2, sub_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    mutate(percent_co=locs/locs_total) %>%
    filter(percent_co>=0.75)
}

## Must have submitted at least 50% of weeks
include_subs <-function(x, df){

  df <-df %>%
    filter(sub_date >= start_date) %>%
    mutate(total=length(unique(sub_date))) %>% 
    group_by(model) %>%
    mutate(subs=length(unique(sub_date))) %>%
    ungroup() %>%
    mutate(percent_sub=subs/total) %>%
    filter(percent_sub>=0.50)

}

############################################################################
############################################################################

load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/updated_cases_nat_state_forecasts_2022-04-04.Rdata")

### National + state submissions
nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  filter(type=="point")

### National + state submissions that met inclusion criteria
teams_nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  include_quant(df=.) %>% #all quantiles present
  include_wks(df=.) %>% #all targets included 
  include_locs(df=.) %>% #at least 50 locations
  include_subs(df=.) %>% #submitted at least 50% of the time
  dplyr::select(model) %>%
  mutate(model_rename=paste0("*", model)) %>%
  distinct()

### Submission plots

nat_st %>%
  dplyr::select(-target, -target_end_date, -value) %>%
  distinct() %>%
  group_by(sub_date, model) %>%
  summarize(locs=n()) %>%
  ungroup() %>%
  mutate(log_loc=log(locs)) %>%
  ungroup() %>%
  mutate(total_sub=length(unique(sub_date))) %>%
  group_by(model) %>%
  mutate(wks_sub=length(sub_date),
         per_sub=wks_sub/total_sub) %>%
  ungroup() %>%
  left_join(., teams_nat_st, by="model") %>%
  mutate(model_rename=ifelse(is.na(model_rename), model, model_rename),
         model_rename=fct_reorder(model_rename, wks_sub)) %>%
  ggplot(., aes(sub_date, model_rename)) +
  geom_tile(aes(fill=log_loc)) +
  scale_fill_gradient(low = "#c7e9c0", high = "#41ab5d") + 
  geom_text(aes(label=locs), size=1.50) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B") +
  theme_bw() +
  labs(title = "A. Primary Model* Submissions for National/State/Territory-Level Case Forecasts Over Time",
       subtitle= "The number of locations forecasted per submission is labled",
       caption = "*Model met inclusion criteria") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        legend.position = "none")

@@

>>>>

---
title: "Supplement 6"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

@@

library(grid)

@@

@@

@@

### Supplement 6

```{r echo=FALSE, warning=F, message=F}

load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/WIS_gee.Rdata") # data
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/model_caseslogstd_noint_gee.rda")

WIS_gee <-WIS_gee %>%
  group_by(model) %>%
  mutate(WIS_log_std=(log(interval_score)-mean(log(interval_score)))/sd(log(interval_score)),
         cases_log_std=(log(true_value)-mean(log(true_value)))/sd(log(true_value))
  ) %>%
  ungroup()

team <-c(unique(as.character(WIS_gee$model))) 
dat_team <-list()
means_sds <-list()
marg_means <-list()

# Model Predictions 
for(i in 1:length(team)){

  dat_team[[i]] <-as.data.frame(WIS_gee %>%
                                  filter(model==team[i]) %>%
                                  mutate(model=droplevels(model)) %>%
                                  ungroup())

  means_sds[[i]] <-dat_team[[i]] %>%
    mutate(mean_log_WIS=mean(log(interval_score)),
           sd_log_WIS=sd(log(interval_score)),
           mean_log_cases=mean(log(true_value)),
           sd_log_cases=sd(log(true_value))
    ) %>%
    dplyr::select(model, contains("mean_"), contains("sd_")) %>%
    distinct()

  marg_means[[i]] <-ggemmeans(model[[i]], terms=c("phase_all", "cases_log_std[-1.96:1.96]")) %>%
    rename(phase_all=x,
           cases=group) %>% 
    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(conf.low*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(conf.high*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           cases=as.numeric(as.character(cases)),
           cases_log_back=(cases*means_sds[[i]]$sd_log_cases) + means_sds[[i]]$mean_log_cases) %>%
    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up),
           cases_back_rescale=exp(cases_log_back))
}
marg_means <-do.call(rbind.data.frame, marg_means)

marg_means <-marg_means %>%

  mutate(model=as.factor(model)) %>%
  mutate(phase_all=as.factor(ifelse(phase_all=="increasing", "Increasing",
                                    ifelse(phase_all=="decreasing", "Decreasing",
                                           ifelse(phase_all=="peak", "Peak", "Nadir")))),
         phase_all=factor(phase_all, 
                          levels = c("Increasing", "Peak", "Decreasing",  "Nadir"))
  )

# Figure
FigS6 <-marg_means %>%
  ggplot(., aes(cases_back_rescale, predicted_back_rescale)) +
  geom_line(aes(color=phase_all)) +
  geom_ribbon(aes(ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale, fill=phase_all), alpha=0.35) +
  scale_fill_manual(breaks = c("Increasing", "Decreasing", "Unclassified", "Peak", "Nadir"), 
                      values=c("#fdae61", "#abdda4", "#bababa", "#d7191c", "#2b83ba")) +
  scale_color_manual(breaks = c("Increasing", "Decreasing", "Unclassified", "Peak", "Nadir"), 
                      values=c("#fdae61", "#abdda4", "#bababa", "#d7191c", "#2b83ba")) +
  labs( y="Marginal Mean \nWeighted Interval Score", x="Reported cases") +
  facet_wrap(~model) +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard,
        legend.title = element_blank(), legend.text=legend.text_standard,
        strip.text=strip.text_standard, strip.background=strip.background_standard,
        axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill=guide_legend(nrow=2, byrow=TRUE))

@@

```{r Fig s6, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=11.5}

grid.draw(shift_legend(FigS6))
